Fractional Types

### 2016-12-19 typespp.tex
tweaks to intro

Both groupoids involve some notion of (semantic) ``equivalence'', which we
would like to capture as first-class entities in a programming language.
In other words, we would like to have some syntactic notion of a type, whose
denotation would be a groupoid with fractional cardinality.  In other words,
a syntactic notion of \emph{fractional types}. Our aim then is to create a
language which possesses ``fractional types'', explore this notion of types,
and its associated operational semantics.

The remainder of the paper is organized as follows. We start by reviewing
necessary background material, consisting of the language $\Pi$ for programming
with isomorphisms or equivalences in a reversible information-preserving way.
Sec.~4 explains the main novel semantic ideas of using $\Pi$ programs to
generate non-trivial groupoids with fractional cardinality, as well as
translating the semantic ideas into an extension of $\Pi$ with new type
constructors denoting non-trivial groupoids and new programs that manipulate
such types.  Sec.~5 presents some example programs.  The last section puts our
work in perspective and concludes.

>>>>

Both groupoids involve some notion of (semantic) ``equivalence'', which we would
like to capture as first-class entities in a programming language.  In other
words, we would like to have some syntactic notion of a type, a \emph{fractional
  type}, whose denotation would be a groupoid with fractional cardinality.  Our
aim then is to create a language which possesses these fractional types, explore
this notion of types, their equivalence, and their associated operational semantics.

The remainder of the paper is organized as follows. We start by reviewing
necessary background material, consisting of the language $\Pi$ for programming
with isomorphisms or equivalences in a reversible information-preserving way.
Sec.~4 explains the main novel semantic ideas of using $\Pi$ programs to
generate non-trivial groupoids with fractional cardinality, as well as
translating the semantic ideas into an extension of $\Pi$ with new type
constructors denoting non-trivial groupoids and new programs that manipulate
such types.  Sec.~5 presents some example programs to illustrate the
expressiveness of the language.  The last section puts our work in perspective
and concludes.
